---
title: "p8105_hw6_xz3078"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
library(patchwork)
```

## Problem 1

### Data Cleaning

```{r}
set.seed(123)

birthweight = read.csv("./birthweight.csv") %>%
  mutate(babysex = as.factor(recode(babysex, "1" = "Male", "2" = "Female")),
         frace = as.factor(recode(frace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other", "9" = "Unknown")),
         malform = as.factor(recode(malform, "0" = "absent", "1" = "present")),
         mrace = as.factor(recode(mrace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other")))

any(is.na(birthweight))
```
There's no missing data in the dataset. 

### Test for relationship between birthweight and each variables

```{r}
par(mfrow = c(3,5))
plot(x = birthweight$bhead, y = birthweight$bwt)
plot(x = birthweight$blength, y = birthweight$bwt)
plot(x = birthweight$delwt, y = birthweight$bwt)
plot(x = birthweight$fincome, y = birthweight$bwt)
plot(x = birthweight$gaweeks, y = birthweight$bwt)
plot(x = birthweight$menarche, y = birthweight$bwt)
plot(x = birthweight$mheight, y = birthweight$bwt)
plot(x = birthweight$momage, y = birthweight$bwt)
plot(x = birthweight$parity, y = birthweight$bwt)
plot(x = birthweight$pnumlbw, y = birthweight$bwt)
plot(x = birthweight$pnumsga, y = birthweight$bwt)
plot(x = birthweight$ppbmi, y = birthweight$bwt)
plot(x = birthweight$ppwt, y = birthweight$bwt)
plot(x = birthweight$smoken, y = birthweight$bwt)
plot(x = birthweight$wtgain, y = birthweight$bwt)
```

```{r}
plot_sex = birthweight %>%
  group_by(babysex) %>%
  ggplot(aes(x = bwt, color = babysex)) +
  geom_density() +
  labs(title = "density plot of bwt by babysex")
plot_mrace = birthweight %>%
  group_by(mrace) %>%
  ggplot(aes(x = bwt, color = mrace)) +
  geom_density() +
  labs(title = "density plot of bwt by mrace")
plot_frace = birthweight %>%
  group_by(frace) %>%
  ggplot(aes(x = bwt, color = frace)) +
  geom_density() +
  labs(title = "density plot of bwt by frace")
plot_malform = birthweight %>%
  group_by(malform) %>%
  ggplot(aes(x = bwt, color = malform)) +
  geom_density() +
  labs(title = "density plot of bwt by malform")

(plot_sex + plot_malform) / (plot_mrace + plot_frace)
```

We can see from the plots that bhead and blength seem to have linear relationships with birthweight. 

### Hypothesis testing

```{r}
anova_1 = anova(lm(bwt ~ bhead + blength, data = birthweight), lm(bwt ~ bhead + blength + babysex, data = birthweight)) %>% 
    broom::tidy() %>% mutate(variable = "babysex")
anova_2 = anova(lm(bwt ~ bhead + blength, data = birthweight), lm(bwt ~ bhead + blength + delwt, data = birthweight)) %>% 
    broom::tidy() %>% mutate(variable = "delwt")
anova_3 = anova(lm(bwt ~ bhead + blength, data = birthweight), lm(bwt ~ bhead + blength + fincome, data = birthweight)) %>% 
    broom::tidy() %>% mutate(variable = "fincome")
anova_4 = anova(lm(bwt ~ bhead + blength, data = birthweight), lm(bwt ~ bhead + blength + frace, data = birthweight)) %>% 
    broom::tidy() %>% mutate(variable = "frace")
anova_5 = anova(lm(bwt ~ bhead + blength, data = birthweight), lm(bwt ~ bhead + blength + gaweeks, data = birthweight)) %>% 
    broom::tidy() %>% mutate(variable = "gaweeks")
anova_6 = anova(lm(bwt ~ bhead + blength, data = birthweight), lm(bwt ~ bhead + blength + malform, data = birthweight)) %>% 
    broom::tidy() %>% mutate(variable = "malform")
anova_7 = anova(lm(bwt ~ bhead + blength, data = birthweight), lm(bwt ~ bhead + blength + menarche, data = birthweight)) %>% 
    broom::tidy() %>% mutate(variable = "menarche")
anova_8 = anova(lm(bwt ~ bhead + blength, data = birthweight), lm(bwt ~ bhead + blength + mheight, data = birthweight)) %>% 
    broom::tidy() %>% mutate(variable = "mheight")
anova_9 = anova(lm(bwt ~ bhead + blength, data = birthweight), lm(bwt ~ bhead + blength + momage, data = birthweight)) %>% 
    broom::tidy() %>% mutate(variable = "momage")
anova_10 = anova(lm(bwt ~ bhead + blength, data = birthweight), lm(bwt ~ bhead + blength + mrace, data = birthweight)) %>% 
    broom::tidy() %>% mutate(variable = "mrace")
anova_11 = anova(lm(bwt ~ bhead + blength, data = birthweight), lm(bwt ~ bhead + blength + parity, data = birthweight)) %>% 
    broom::tidy() %>% mutate(variable = "parity")
anova_12 = anova(lm(bwt ~ bhead + blength, data = birthweight), lm(bwt ~ bhead + blength + pnumlbw, data = birthweight)) %>% 
    broom::tidy() %>% mutate(variable = "pnumlbw")
anova_13 = anova(lm(bwt ~ bhead + blength, data = birthweight), lm(bwt ~ bhead + blength + pnumsga, data = birthweight)) %>% 
    broom::tidy() %>% mutate(variable = "pnumsga")
anova_14 = anova(lm(bwt ~ bhead + blength, data = birthweight), lm(bwt ~ bhead + blength + ppbmi, data = birthweight)) %>% 
    broom::tidy() %>% mutate(variable = "ppbmi")
anova_15 = anova(lm(bwt ~ bhead + blength, data = birthweight), lm(bwt ~ bhead + blength + ppwt, data = birthweight)) %>% 
    broom::tidy() %>% mutate(variable = "ppwt")
anova_16 = anova(lm(bwt ~ bhead + blength, data = birthweight), lm(bwt ~ bhead + blength + smoken, data = birthweight)) %>% 
    broom::tidy() %>% mutate(variable = "smoken")
anova_17 = anova(lm(bwt ~ bhead + blength, data = birthweight), lm(bwt ~ bhead + blength + wtgain, data = birthweight)) %>% 
    broom::tidy() %>% mutate(variable = "wtgain")
rbind(anova_1, anova_2, anova_3, anova_4, anova_5, anova_6, anova_7, anova_8, anova_9, anova_10, anova_11, anova_12, anova_13, anova_14, anova_15, anova_16, anova_17) %>% drop_na(df)
```

From the anova test above, we can see the bhead, blength, babysex, delwt, fincome, frace, gaweeks, mheight, momage, mrace, ppwt, wtgain are all related to birthweight since their p value are extremely small. 

### Plot of model residuals against fitted values

```{r}
fit = lm(bwt ~ bhead + blength + babysex + delwt + fincome + frace + gaweeks + mheight + momage + mrace + ppwt + wtgain, data = birthweight)
birthweight %>% 
  modelr::add_predictions(fit) %>%
  modelr::add_residuals(fit) %>%
  ggplot(aes(x = pred, y = resid)) +
  geom_point() 
```

To build my model, I first find the variables that have clear linear relationship with birthweight and then do the hypothesis test to find other potential significant variables. Finally, I put all variables in one model. 

### Compare models

```{r warning = FALSE}
cv_df = 
  crossv_mc(birthweight, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)) %>%
  mutate(
    mod_1 = map(train, ~lm(bwt ~ bhead + blength + babysex + delwt + fincome + frace + gaweeks + mheight + momage + mrace + ppwt + wtgain, data = .x)),
    mod_2 = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    mod_3 = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead * blength +bhead * babysex + babysex * blength + babysex * blength * bhead, data = .x))) %>% 
  mutate(
    rmse_mod_1 = map2_dbl(mod_1, test, ~rmse(model = .x, data = .y)),
    rmse_mod_2 = map2_dbl(mod_2, test, ~rmse(model = .x, data = .y)),
    rmse_mod_3 = map2_dbl(mod_3, test, ~rmse(model = .x, data = .y)))

cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

From the violin plot above, we can see that model 1 has the least rmse among all three models. This means that the model I built is better compare with the other two. 

## Problem 2

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

### r^2

```{r}
boot_sample = function(df) {
  sample_frac(df, replace = TRUE)
}

boot_straps = 
  data_frame(
    strap_number = 1:5000,
    strap_sample = rerun(5000, boot_sample(weather_df))
  )

bootstrap_results_r = 
  boot_straps %>% 
  mutate(
    models = map(strap_sample, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::glance)) %>% 
  select(-strap_sample, -models) %>% 
  unnest(results) 

bootstrap_results_r %>% 
  summarize(
    ci_lower = quantile(adj.r.squared, 0.025), 
    ci_upper = quantile(adj.r.squared, 0.975)) %>% 
  knitr::kable(digits = 3)

bootstrap_results_r %>% 
  ggplot(aes(x = adj.r.squared)) + 
  geom_density() +
  labs(title = "distribution of r^2")
```

The distribution of r^2 is approximately normal where the center is around 0.91 with a 95% confidence interval of (0.894, 0.927). 

### log(beta_0*beta_1)

```{r}
bootstrap_results_log = 
  boot_straps %>% 
  mutate(
    models = map(strap_sample, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::tidy)) %>% 
  select(-strap_sample, -models) %>% 
  unnest(results) %>%
  select(term, estimate) %>%
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) %>%
  unnest() %>%
  rename(beta_0 = "(Intercept)", beta_1 = tmin) %>%
  mutate(log = log(beta_0 * beta_1))

bootstrap_results_log %>% 
  summarize(
    ci_lower = quantile(log, 0.025), 
    ci_upper = quantile(log, 0.975)) %>% 
  knitr::kable(digits = 3)

bootstrap_results_log %>% 
  ggplot(aes(x = log)) + 
  geom_density() +
  labs(x = "log(beta_0*beta_1)", title = "distribution of log(beta_0*beta_1)")
```

The distribution of log(beta_0*beta_1) is approximately normal where the center is around 2.02 with a 95% confidence interval of (1.964, 2.058). 

